---
title: "Coronavirus in the Netherlands"
author: "Piet Stam"
# runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    # social: ["facebook", "twitter", "linkedin"]
    source_code: embed
    vertical_layout: fill
---

```{r setup, include=FALSE}
#------------------ Packages ------------------
library(flexdashboard)
# install.packages("devtools")
# devtools::install_github("RamiKrispin/coronavirus", force = TRUE)
library(coronavirus)
data(coronavirus)
update_dataset()
# View(coronavirus)
# max(coronavirus$date)

`%>%` <- magrittr::`%>%`
#------------------ Parameters ------------------
# Set colors
# https://www.w3.org/TR/css-color-3/#svg-color
confirmed_color <- "purple"
active_color <- "#1f77b4"
recovered_color <- "forestgreen"
death_color <- "red"
#------------------ Data ------------------
# Note that we exclude the Dutch Carribean islands 
df <- coronavirus %>%
  # dplyr::filter(date == max(date)) %>%
  dplyr::filter(country == "Netherlands") %>%
  dplyr::filter(province == "") %>%
  dplyr::group_by(country, type) %>%
  dplyr::summarise(total = sum(cases)) %>%
  tidyr::pivot_wider(
    names_from = type,
    values_from = total
  ) %>%
  # dplyr::mutate(unrecovered = confirmed - ifelse(is.na(recovered), 0, recovered) - ifelse(is.na(death), 0, death)) %>%
  dplyr::mutate(unrecovered = confirmed - ifelse(is.na(death), 0, death)) %>%
  dplyr::arrange(-confirmed) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(country = dplyr::if_else(country == "United Arab Emirates", "UAE", country)) %>%
  dplyr::mutate(country = dplyr::if_else(country == "Mainland China", "China", country)) %>%
  dplyr::mutate(country = dplyr::if_else(country == "North Macedonia", "N.Macedonia", country)) %>%
  dplyr::mutate(country = trimws(country)) %>%
  dplyr::mutate(country = factor(country, levels = country))

# Note that we exclude the Dutch Carribean islands 
df_daily <- coronavirus %>%
  dplyr::filter(country == "Netherlands") %>%
  dplyr::filter(province == "") %>%
  dplyr::filter(date >= "2020-02-27") %>%
  dplyr::group_by(date, type) %>%
  dplyr::summarise(total = sum(cases, na.rm = TRUE)) %>%
  tidyr::pivot_wider(
    names_from = type,
    values_from = total
  ) %>%
  dplyr::arrange(date) %>%
  dplyr::ungroup() %>%
  #dplyr::mutate(active = confirmed - death - recovered) %>%
  dplyr::mutate(active = confirmed - death) %>%
  dplyr::mutate(
    confirmed_cum = cumsum(confirmed),
    death_cum = cumsum(death),
    # recovered_cum = cumsum(recovered),
    active_cum = cumsum(active)
  )


df1 <- coronavirus %>% dplyr::filter(date == max(date))
```

Country comparison
=======================================================================


Column {data-width=400}
-------------------------------------


### **Cumulative number of deaths** (per million inhabitants, data of `r format(max(coronavirus$date), "%A %B %d, %Y")`)
    
```{r}
daily_death <- coronavirus %>%
  dplyr::filter(type == "death") %>%
  dplyr::filter(date >= "2020-02-27") %>%
  dplyr::mutate(country = country) %>%
  dplyr::group_by(date, country) %>%
  dplyr::summarise(total = sum(cases)) %>%
  dplyr::group_by(country) %>% 
  dplyr::mutate(total = cumsum(total)) %>% 
  dplyr::ungroup() %>%
  tidyr::pivot_wider(names_from = country, values_from = total)

#----------------------------------------
# Plotting the data

daily_death %>%
  plotly::plot_ly() %>%
  plotly::add_trace(
    x = ~date,
    y = ~Germany*1000000/83149300,
    type = "scatter",
    mode = "lines+markers",
    name = "Germany"
  ) %>%
  plotly::add_trace(
    x = ~date,
    y = ~Netherlands*1000000/17452431,
    type = "scatter",
    mode = "lines+markers",
    name = "Netherlands"
  ) %>%
  plotly::add_trace(
    x = ~date,
    y = ~Italy*1000000/60238522,
    type = "scatter",
    mode = "lines+markers",
    name = "Italy"
  ) %>%
  plotly::add_trace(
    x = ~date,
    y = ~Spain*1000000/47100396,
    type = "scatter",
    mode = "lines+markers",
    name = "Spain"
  ) %>%
  plotly::add_trace(
    x = ~date,
    y = ~Belgium*1000000/11524454,
    type = "scatter",
    mode = "lines+markers",
    name = "Belgium"
  ) %>%
  plotly::layout(
    title = "",
    legend = list(x = 0.05, y = 0.95),
    yaxis = list(title = "Semilogarithmic scale", hoverformat = ',.0f', tickformat = ',.0f', type = "log", tick0 = 1, dtick = 1),
    xaxis = list(title = "Date"),
    # paper_bgcolor = "black",
    # plot_bgcolor = "black",
    # font = list(color = 'white'),
    hovermode = "compare",
    margin = list(
      # l = 60,
      # r = 40,
      b = 10,
      t = 10,
      pad = 2
    )
  )
```

### **Cases distribution by type** (absolute numbers, data of `r format(max(coronavirus$date), "%A %B %d, %Y")`)

```{r daily_summary}
df_EU <- coronavirus %>%
  # dplyr::filter(date == max(date)) %>%
  dplyr::filter(country == "Netherlands" |
    country == "Germany" |
    country == "Spain" |
    country == "Italy" |
    country == "Belgium") %>%
  dplyr::group_by(country, type) %>%
  dplyr::summarise(total = sum(cases)) %>%
  tidyr::pivot_wider(
    names_from = type,
    values_from = total
  ) %>%
  # dplyr::mutate(unrecovered = confirmed - ifelse(is.na(recovered), 0, recovered) - ifelse(is.na(death), 0, death)) %>%
  dplyr::mutate(unrecovered = confirmed - ifelse(is.na(death), 0, death)) %>%
  dplyr::arrange(confirmed) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(country = dplyr::if_else(country == "United Arab Emirates", "UAE", country)) %>%
  dplyr::mutate(country = dplyr::if_else(country == "Mainland China", "China", country)) %>%
  dplyr::mutate(country = dplyr::if_else(country == "North Macedonia", "N.Macedonia", country)) %>%
  dplyr::mutate(country = trimws(country)) %>%
  dplyr::mutate(country = factor(country, levels = country))

plotly::plot_ly(
  data = df_EU,
  x = ~country,
  # y = ~unrecovered,
  y = ~ confirmed,
  # text =  ~ confirmed,
  # textposition = 'auto',
  type = "bar",
  name = "Confirmed",
  marker = list(color = active_color)
) %>%
  plotly::add_trace(
    y = ~death,
    # text =  ~ death,
    # textposition = 'auto',
    name = "Death",
    marker = list(color = death_color)
  ) %>%
  plotly::layout(
    barmode = "stack",
    yaxis = list(title = "Total cases", hoverformat = ',.0f', tickformat = ',.0f'),
    xaxis = list(title = ""),
    hovermode = "compare",
    margin = list(
      # l = 60,
      # r = 40,
      b = 10,
      t = 10,
      pad = 2
    )
  )
```


Netherlands
=======================================================================

Row {data-width=400}
-----------------------------------------------------------------------

### confirmed {.value-box}

```{r}

valueBox(
  value = paste(format(sum(df$confirmed), big.mark = ","), "", sep = " "),
  caption = "Total confirmed cases",
  icon = "fas fa-user-md",
  color = confirmed_color
)
```


<!-- ### active {.value-box} -->

<!-- ```{r} -->
<!-- valueBox( -->
<!--   value = paste(format(sum(df$unrecovered, na.rm = TRUE), big.mark = ","), " (", -->
<!--     round(100 * sum(df$unrecovered, na.rm = TRUE) / sum(df$confirmed), 1), -->
<!--     "%)", -->
<!--     sep = "" -->
<!--   ), -->
<!--   caption = "Active cases (% of total cases)", icon = "fas fa-ambulance", -->
<!--   color = active_color -->
<!-- ) -->
<!-- ``` -->

### death {.value-box}

```{r}

valueBox(
  value = paste(format(sum(df$death, na.rm = TRUE), big.mark = ","), " (",
    round(100 * sum(df$death, na.rm = TRUE) / sum(df$confirmed), 1),
    "%)",
    sep = ""
  ),
  caption = "Death cases (death rate)",
  icon = "fas fa-heart-broken",
  color = death_color
)
```


Row
-----------------------------------------------------------------------

### **Cumulative cases by type** (Netherlands only, data of `r format(max(coronavirus$date), "%A %B %d, %Y")`)
    
```{r}
plotly::plot_ly(data = df_daily) %>%
  plotly::add_trace(
    x = ~date,
    # y = ~active_cum,
    y = ~confirmed_cum,
    type = "scatter",
    mode = "lines+markers",
    # name = "Active",
    name = "Confirmed",
    line = list(color = active_color),
    marker = list(color = active_color)
  ) %>%
  plotly::add_trace(
    x = ~date,
    y = ~death_cum,
    type = "scatter",
    mode = "lines+markers",
    name = "Death",
    line = list(color = death_color),
    marker = list(color = death_color)
  ) %>%
#  plotly::add_annotations(
#    x = as.Date("2020-02-27"),
#    y = 0,
#    text = paste("First case"),
#    xref = "x",
#    yref = "y",
#    arrowhead = 5,
#    arrowhead = 3,
#    arrowsize = 1,
#    showarrow = TRUE,
#    ax = -10,
#    ay = -90
#  ) %>%
#  plotly::add_annotations(
#    x = as.Date("2020-03-06"),
#    y = 0,
#    text = paste("First death"),
#    xref = "x",
#    yref = "y",
#    arrowhead = 5,
#    arrowhead = 3,
#    arrowsize = 1,
#    showarrow = TRUE,
#    ax = -80,
#    ay = -45
#  ) %>%
  plotly::add_annotations(
    x = as.Date("2020-03-06"),
    y = 0,
    text = paste(
      "Stay at home",
      "<BR>",
      "Brabant"
    ),
    xref = "x",
    yref = "y",
    arrowhead = 5,
    arrowhead = 3,
    arrowsize = 1,
    showarrow = TRUE,
    ax = -45,
    ay = -90
  ) %>%
  plotly::add_annotations(
    x = as.Date("2020-03-09"),
    y = 0,
    text = paste(
      "No handshakes"
    ),
    xref = "x",
    yref = "y",
    arrowhead = 5,
    arrowhead = 3,
    arrowsize = 1,
    showarrow = TRUE,
    ax = -10,
    ay = -45
  ) %>%
  plotly::add_annotations(
    x = as.Date("2020-03-12"),
    y = 0,
    text = paste(
      "Stay at home",
      "<BR>",
      "whole country"
    ),
    xref = "x",
    yref = "y",
    arrowhead = 5,
    arrowhead = 3,
    arrowsize = 1,
    showarrow = TRUE,
    ax = -10,
    ay = -90
  )  %>%
  plotly::add_annotations(
    x = as.Date("2020-03-15"),
    y = 0,
    text = paste(
      "April 6th",
      "<br>",
      "measure"
    ),
    xref = "x",
    yref = "y",
    arrowhead = 5,
    arrowhead = 3,
    arrowsize = 1,
    showarrow = TRUE,
    ax = 20,
    ay = -110
  ) %>%
  plotly::add_annotations(
    x = as.Date("2020-03-19"),
    y = 0,
    text = paste(
      "Nursing homes",
      "<br>",
      "closed"
    ),
    xref = "x",
    yref = "y",
    arrowhead = 5,
    arrowhead = 3,
    arrowsize = 1,
    showarrow = TRUE,
    ax = 20,
    ay = -45
  )  %>%
  plotly::add_annotations(
    x = as.Date("2020-03-31"),
    y = 0,
    text = paste(
      "April 28th",
      "<br>",
      "measure"
    ),
    xref = "x",
    yref = "y",
    arrowhead = 5,
    arrowhead = 3,
    arrowsize = 1,
    showarrow = TRUE,
    ax = 0,
    ay = -90
  ) %>%
  plotly::add_annotations(
    x = as.Date("2020-04-21"),
    y = 0,
    text = paste(
      "Extended",
      "<br>",
      "May 11th",
      "<br>",
      "measure"
    ),
    xref = "x",
    yref = "y",
    arrowhead = 5,
    arrowhead = 3,
    arrowsize = 1,
    showarrow = TRUE,
    ax = 0,
    ay = -90
  ) %>%
  plotly::layout(
    title = "",
    yaxis = list(title = "Semilogarithmic scale", type = "log", tick0 = 1, dtick = 1, hoverformat = ',.0f', tickformat = ',.0f'),
    xaxis = list(title = "Date"),
    legend = list(x = 0.05, y = 0.95),
    hovermode = "compare"
  )
```

About
=======================================================================

**The Coronavirus Dashboard: the case of Netherlands**

[The Coronavirus Dashboard: the case of the
Netherlands](https://pietstam.nl/coronavirus-dashboard/)
provides an overview of the 2019 Novel Coronavirus COVID-19 (2019-nCoV)
epidemic for the Netherlands. It is built with R using the R Markdown
framework and is based on the
[dashboard](https://www.antoinesoetewey.com/files/coronavirus-dashboard.html)
by Antoine Soetewey, who uses data from Belgium. His dashboard is an
adaptation of the original
[dashboard](https://ramikrispin.github.io/coronavirus_dashboard/) by
Rami Krispin.

**New features**

[2020-04-27] We changed the scales of two out of three figures to a logarithmic one. In addition some formatting was done.

[2020-04-06] We compare the number of deaths instead of the number of infected cases on the comparison tab of the dashboard. The reason is that we assume that bias in the country comparison because of measurement errors is smaller in this case.

[2020-03-29] We denoted the numbers in the left figure on the Comparison tab of the dashboard in terms of 1 million inhabitants living in the respective countries. The total number of inhabitants are copied from the [Wikipedia](https://en.wikipedia.org/wiki/Main_Page) [List of countries and dependencies by population](https://en.wikipedia.org/wiki/List_of_countries_and_dependencies_by_population) (visited at 2020-04-06).

**Code**

The code behind this dashboard is available on [GitHub](https://github.com/pjastam/coronavirus-dashboard){target="_blank"}.

**Data**

The input data for this dashboard is from the [`{coronavirus}`](https://github.com/RamiKrispin/coronavirus){target="_blank"} R package. Make sure to download the development version of the package to have the latest data:

```
install.packages("devtools")
devtools::install_github("RamiKrispin/coronavirus")
```

The data and dashboard are refreshed on a daily basis.

The raw data is pulled from the Johns Hopkins University Center for Systems Science and Engineering (JHU CCSE) Coronavirus [repository](https://github.com/RamiKrispin/coronavirus-csv){target="_blank"}.

**Contact**

For any question or feedback, you can either open an <a href="https://github.com/pjastam/coronavirus-dashboard/issues">issue</a> or contact me on <a href="https://twitter.com/pjastam">Twitter</a>. More information about this dashboard can be found in this [article](https://www.statsandr.com/blog/how-to-create-a-simple-coronavirus-dashboard-specific-to-your-country-in-r/).

**Update**

The data is as of `r format(max(coronavirus$date), "%A %B %d, %Y")` and the dashboard has been updated on `r format(Sys.time(), "%A %B %d, %Y")`.
